---
title: "`r gsub('.Rmd', ' ', gsub('_', ' ', knitr::current_input()))`"
author: "Holly Beale"
date: "`r format(Sys.time(), '%B %d, %Y')`"
format: 
  gfm:
    df-print: kable
---

```{r setup}

library(tidyverse)
library(here)
library(janitor)

```


```{r}
gdc_sample_sheet <- read_tsv(here("2025-05_tcga_luad_sig_from_bam/gdc_sample_sheet.2025-05-22.tsv"))
gdc_manifest <- read_tsv(here("2025-05_tcga_luad_sig_from_bam/gdc_manifest.2025-05-22.132704.txt"))
```

# Compare manifest and sample sheet contents

```{r}

identical(gdc_sample_sheet$`File ID`, gdc_manifest$id)


```
They are the same


# review sample features
```{r}

n_unique <- function(x) length(unique(x))

unique_sample_vals <- gdc_sample_sheet %>%
  summarize(across(everything(),  n_unique)) %>%
  pivot_longer(everything())

unique_sample_vals
  
```

## What are multi-value columns?
Notes:
anything with "1" or 541 is uninteresting


```{r}

table(gdc_sample_sheet$`Tumor Descriptor`)
table(gdc_sample_sheet$`Specimen Type`)
table(gdc_sample_sheet$`Preservation Method`)

```

# fix column names
```{r}

lc_no_space <- function(x) str_replace_all(x, " ", "_") %>% tolower()
  

gdc_sample_sheet_renamed <- gdc_sample_sheet %>% rename_with(lc_no_space)
```

# review duplicate sample IDs

```{r}

dupe_sample_ids <- gdc_sample_sheet_renamed$sample_id[duplicated(gdc_sample_sheet_renamed$sample_id)]


gdc_sample_sheet_renamed %>%
  filter(sample_id %in% dupe_sample_ids)
```

# review duplicate case IDs

```{r}

dupe_case_ids <- gdc_sample_sheet_renamed$case_id[duplicated(gdc_sample_sheet_renamed$case_id)]



gdc_sample_sheet_renamed %>%
  filter(case_id %in% dupe_case_ids,
         #! sample_id %in% dupe_sample_ids
         ) %>%
  arrange(case_id)



```

 
Create bam manifest 
```{r}

gdc_sample_sheet_renamed_for_bam_manifest <- gdc_sample_sheet_renamed %>%
  mutate(tcga_id = paste(sample_id, file_id),
         file_location = paste0("/mnt/data/tcga/", file_id, "/", file_name),
         feature1 = lc_no_space(preservation_method),
         feature2 = lc_no_space(specimen_type))
  
gdc_sample_sheet_renamed_for_bam_manifest %>%
  select(tcga_id,
         file_location,
         feature1,
         feature2) %>%
  write_tsv(paste0("bam_manifest.tcga.", 
                   nrow(gdc_sample_sheet_renamed_for_bam_manifest), 
                   "_bam_files.tsv"),
            col_names = FALSE)

```
 
 
sync with luad_s34f_wt_manifest.tsv

```{r}
luad_u2af1_s34f_wt_manifest <- read_tsv(here("2025-05_tcga_luad_sig_from_bam/luad_u2af1_s34f_wt_manifest.tsv"),
                                        col_names = c("id", "mutation"))

gdc_sample_sheet_renamed_with_mutations <- 
  gdc_sample_sheet_renamed %>%
  mutate(mutation_match_code = str_remove(file_name, ".rna_seq.genomic.gdc_realn.bam")) %>%
  left_join(luad_u2af1_s34f_wt_manifest,
            by=c("mutation_match_code" = "id")) 

gdc_sample_sheet_renamed_with_mutations %>%
  tabyl(mutation)

```


```{r}
gdc_sample_sheet_renamed_with_mutations %>%
  filter(is.na(mutation))
```

 # select all u2af1-s34f samples
```{r}


u2af1_s34f_sample_sheet <- gdc_sample_sheet_renamed_with_mutations %>%
  filter(! is.na(mutation)) %>%
  filter(mutation == "u2af1-s34f")

nrow(u2af1_s34f_sample_sheet)

length(unique(u2af1_s34f_sample_sheet$case_id))

cat (u2af1_s34f_sample_sheet$file_id)

```
 
 
 new manifest
 
 # Exclude samples without mutation info
 randomly select duplicates 
 include all u2af1-s34f samples
 check if u2af1 mutation status disagrees in publication
 
```{r}

set.seed(124343)
u2af1_wt_samples_one_per_case <- gdc_sample_sheet_renamed_with_mutations %>%
  filter(! is.na(mutation)) %>%
  filter(preservation_method == "Unknown",
         tumor_descriptor == "Primary",
         mutation != "u2af1-s34f") %>%
  group_by(case_id) %>%
   slice_sample(n=1)



```
 
```{r}

gdc_manifest_50_samples <- read_tsv(here("2025-05_tcga_luad_sig_from_bam/gdc_manifest.50_samples.2025-05-22.132704.txt"))

new_u2af1_wt_samples_one_per_case <- u2af1_wt_samples_one_per_case %>%
  filter(! file_id %in% gdc_manifest_50_samples$id)

set.seed(124343)
u2af1_wt_samples_for_batch2 <- new_u2af1_wt_samples_one_per_case %>%
  ungroup %>%
  slice_sample(n=(50-12))


```
 
next 50 
```{r}


batch2_manifest <- gdc_manifest %>%
  filter(id %in% 
           c(u2af1_s34f_sample_sheet$file_id,
             u2af1_wt_samples_for_batch2$file_id))

batch2_manifest %>%
  write_tsv(here("2025-05_tcga_luad_sig_from_bam/gdc_manifest.50_samples.batch2.2025-05-28.txt"))
  

```

# show OCT samples
```{r}


gdc_sample_sheet_renamed_with_mutations %>% 
  filter(preservation_method == "OCT", 
           file_id %in% batch2_manifest$id)
```

```{r}


# this is the sample that wasn't detected when i repeated the analysis with train/test datasets
# TCGA-55-7727-01A_86c05b02-68d0-473d-8aea-ab501cb40d29

gdc_sample_sheet_renamed_with_mutations %>%
  filter(str_detect(case_id, "TCGA-55-7727"))


```
 
 