---
title: "`r gsub('.Rmd', ' ', gsub('_', ' ', knitr::current_input()))`"
author: "Holly Beale"
date: "`r format(Sys.time(), '%B %d, %Y')`"
format: 
  gfm:
    df-print: kable
---

#load_libraries

```{r setup}

library(tidyverse)
library(janitor)
```
# Previous findings for comparison
from splicedice manuscript draft:
Of the 48,837 samples that we tested, only 22 were significant matches to the U2AF1-S34F signature.

# Load data
```{r}
results_file <- "/mnt/output/splicedice_U2AF1_sig_in_intropolis_2025.04.29_23.43.54//U2AF1_sig_in_intropolis.pvals.tsv"

results <- read_tsv(results_file)
```

data features
```{r}

system(paste("md5sum", results_file), intern = TRUE)

dim(results)


```

The number of samples in this results file:
`r prettyNum(ncol(results)-1, big.mark=",",scientific=FALSE)`

# pivot for managability
```{r}

results_long <- results %>%
  pivot_longer (-query, names_to = "SRA_dataset", values_to = "p_value") %>%
  mutate(neglog10p = -log10(p_value),
         basic_significance = p_value < 0.05,
         bonf_significance = p_value/(ncol(results)-1) < -0.05)

```
# Exclude NA data
```{r}
results_long_NA_annotated <- results_long %>%
  group_by(SRA_dataset) %>%
  mutate(two_NAs = all(is.na(p_value))) 

results_long_NA_annotated %>%
  filter(two_NAs) %>%
  arrange(SRA_dataset) %>%
  head

results_long_NA_excluded <- results_long_NA_annotated %>%
  filter(!two_NAs)


```




# survey of results
```{r}
summary(results_long_NA_excluded$p_value)
```


```{r}
tabyl(results_long_NA_excluded,
      query,
      basic_significance)
```


```{r}
tabyl(results_long_NA_excluded,
      query,
      basic_significance) %>%
  adorn_percentages() %>%
  adorn_pct_formatting(digits = 2)
```

## bonf_significance
I calculated bonf_significance in case the p values were not already adjusted for multiple testing; no values exceed it
```{r}
tabyl(results_long_NA_excluded,
                 bonf_significance)
```

# Show the significant results
```{r}

results_long_NA_excluded %>%
  filter(query == "u2af1-s34f_over_u2af1-wt",
         basic_significance)

```

All significant results have the same p-value. Weird. Look at all samples with that p-value
26 additional samples have that p-value. Those are wt/mutant, not mutant/wt

```{r}

results_long_NA_excluded %>%
  filter(p_value > 0.02476730,
         p_value < 0.02476732)

```


# visualize distribution of p values
histogram

```{r}

ggplot(results_long_NA_excluded) + 
  geom_histogram(aes(x=p_value, fill = basic_significance)) +
  scale_fill_brewer(palette = "Set1") +
  facet_wrap(~query, ncol = 1)

```

```{r}

ggplot(results_long_NA_excluded) + 
  geom_boxplot(aes(x=p_value))  +
  facet_wrap(~query, ncol = 1)

```
# Review -log10(p value) 
boxplot
```{r}
ggplot(results_long_NA_excluded) + 
  geom_boxplot(aes(x=neglog10p)) +
  scale_fill_brewer(palette = "Set1")
```

histogram
```{r}
ggplot(results_long_NA_excluded) + 
  geom_histogram(aes(x=neglog10p, fill = basic_significance)) +
  scale_fill_brewer(palette = "Set1") +
  facet_wrap(~query, ncol = 1)

```

Conclusion
Here we find 10 significant matches to the u2af1 signature (assuming a p-value < 0.05 in the u2af1-s34f_over_u2af1-wt query means the same thing Dennis means when he says "significant matches to the U2AF1-S34F signature"). 

If we want to pursue this, if we can find the key to connect intropolis IDs (e.g. 1-48000) to SRA IDs, we could check if we see what Dennis saw, the u2af1-s34f mutation in the SRA data.