---
title: "`r gsub('.Rmd', ' ', gsub('_', ' ', knitr::current_input()))`"
author: "Holly Beale"
date: "`r format(Sys.time(), '%B %d, %Y')`"
format: 
  gfm:
    df-print: kable
---

```{r setup}

library(tidyverse)
library(here)
library(janitor)

```


```{r}
gdc_sample_sheet <- read_tsv(here("2025-05_tcga_luad_download/gdc_sample_sheet.2025-05-22.tsv"))
gdc_manifest <- read_tsv(here("2025-05_tcga_luad_download/gdc_manifest.2025-05-22.132704.txt"))
```

# Compare manifest and sample sheet contents

```{r}

identical(gdc_sample_sheet$`File ID`, gdc_manifest$id)


```
They are the same


# review sample features
```{r}

n_unique <- function(x) length(unique(x))

unique_sample_vals <- gdc_sample_sheet %>%
  summarize(across(everything(),  n_unique)) %>%
  pivot_longer(everything())

unique_sample_vals
  
```

## What are multi-value columns?
Notes:
anything with "1" or 541 is uninteresting


```{r}

table(gdc_sample_sheet$`Tumor Descriptor`)
table(gdc_sample_sheet$`Specimen Type`)
table(gdc_sample_sheet$`Preservation Method`)

```

# fix column names
```{r}

lc_no_space <- function(x) str_replace_all(x, " ", "_") %>% tolower()
  

gdc_sample_sheet_renamed <- gdc_sample_sheet %>% rename_with(lc_no_space)
```

# review duplicate sample IDs

```{r}

dupe_sample_ids <- gdc_sample_sheet_renamed$sample_id[duplicated(gdc_sample_sheet_renamed$sample_id)]


gdc_sample_sheet_renamed %>%
  filter(sample_id %in% dupe_sample_ids)
```

# review duplicate case IDs

```{r}

dupe_case_ids <- gdc_sample_sheet_renamed$case_id[duplicated(gdc_sample_sheet_renamed$case_id)]



gdc_sample_sheet_renamed %>%
  filter(case_id %in% dupe_case_ids,
         #! sample_id %in% dupe_sample_ids
         ) %>%
  arrange(case_id)



```

 
Create bam manifest 
```{r}

gdc_sample_sheet_renamed_for_bam_manifest <- gdc_sample_sheet_renamed %>%
  mutate(tcga_id = paste(sample_id, file_id),
         file_location = paste0("/mnt/data/tcga/", file_id, "/", file_name),
         feature1 = lc_no_space(preservation_method),
         feature2 = lc_no_space(specimen_type))
  
gdc_sample_sheet_renamed_for_bam_manifest %>%
  select(tcga_id,
         file_location,
         feature1,
         feature2) %>%
  write_tsv(paste0("bam_manifest.tcga.", 
                   nrow(gdc_sample_sheet_renamed_for_bam_manifest), 
                   "_bam_files.tsv"),
            col_names = FALSE)

```
 
 
sync with luad_s34f_wt_manifest.tsv

```{r}
luad_u2af1_s34f_wt_manifest <- read_tsv(here("2025-05_tcga_luad_download/luad_u2af1_s34f_wt_manifest.tsv"),
                                        col_names = c("id", "mutation"))

gdc_sample_sheet_renamed_with_mutations <- 
  gdc_sample_sheet_renamed %>%
  mutate(mutation_match_code = str_remove(file_name, ".rna_seq.genomic.gdc_realn.bam")) %>%
  left_join(luad_u2af1_s34f_wt_manifest,
            by=c("mutation_match_code" = "id")) 

gdc_sample_sheet_renamed_with_mutations %>%
  tabyl(mutation)

```


```{r}
gdc_sample_sheet_renamed_with_mutations %>%
  filter(is.na(mutation))
```

 # Exclude samples without mutation info
 randomly select duplicates 
 include all u2af1-s34f samples
 check if u2af1 mutation status disagrees in publication
 
 
```{r}


u2af1_s34f_sample_sheet <- gdc_sample_sheet_renamed_with_mutations %>%
  filter(! is.na(mutation)) %>%
  filter(mutation == "u2af1-s34f")

nrow(u2af1_s34f_sample_sheet)

length(unique(u2af1_s34f_sample_sheet$case_id))


  
```
 
 