---
title: "`r gsub('.Rmd', ' ', gsub('_', ' ', knitr::current_input()))`"
author: "Holly Beale"
date: "`r format(Sys.time(), '%B %d, %Y')`"
format: 
  gfm:
    df-print: kable
---

```{r setup}

library(tidyverse)
library(here)
library(janitor)

```

```{r}
LUAD_soulette <- read_tsv("LUAD_601_RNA_summary_Final495SamplesForAnalysis.tsv")  %>%
  mutate(sample_id_key = str_extract(barcode, "^TCGA-\\w+-\\w+-\\w+"))
```
```{r}
head(LUAD_soulette)
nrow(LUAD_soulette)
```


```{r}
gdc_sample_sheet <- read_tsv(here("2025-05_tcga_luad_sig_from_bam/gdc_sample_sheet.2025-05-22.tsv"))
gdc_manifest <- read_tsv(here("2025-05_tcga_luad_sig_from_bam/gdc_manifest.2025-05-22.132704.txt"))

nrow(gdc_sample_sheet)
nrow(gdc_manifest)

gdc_sample_sheet_renamed <- gdc_sample_sheet %>% clean_names() 

```

# Compare manifest and sample sheet contents

```{r}

identical(gdc_sample_sheet$`File ID`, gdc_manifest$id)

n_distinct(gdc_sample_sheet$`File ID`)


```
They are the same


# review sample features
```{r}

n_unique <- function(x) length(unique(x))

unique_sample_vals <- LUAD_soulette %>%
  summarize(across(everything(),  n_unique)) %>%
  pivot_longer(everything())

unique_sample_vals %>%
  filter(! value == 1,
         ! value == nrow(LUAD_soulette))
  
```

## What are multi-value columns?
Notes:
anything with "1" or 541 is uninteresting


```{r}

LUAD_soulette %>%
  select(unique_sample_vals %>% 
           filter(value > min(value),
                  value < max(value)) %>%
           pull(name)) %>%
  apply(2, table)
  

# head(LUAD_soulette$reason)
# head(LUAD_soulette$state)
```

# check overlap with gdc sample sheet
```{r}

sample_sheet_original <- read_tsv("gdc_sample_sheet.2025-05-22.tsv")

sample_sheet <- clean_names(sample_sheet_original)

sample_sheet_colnames <- tibble(original = colnames(sample_sheet_original),
                                groomed = colnames(sample_sheet))




```

# Figure out which columns match
```{r}

id_from_gdc_sample_sheet_file_name_column <- "325c7d6b-292d-46e5-85d1-785ae8a48c33"

LUAD_soulette %>%
  filter(str_detect(filename, id_from_gdc_sample_sheet_file_name_column))

LUAD_soulette %>%
  filter(str_detect(analysis_id, id_from_gdc_sample_sheet_file_name_column))

```
```{r}

# Count occurrences of each sample_id_key
LUAD_soulette %>%
  count(sample_id_key, name = "n_rows") %>%
  filter(n_rows > 1) %>%
  arrange(desc(n_rows))

joined_data <- left_join(LUAD_soulette, sample_sheet,
          by=c("sample_id_key"="sample_id")) 

dupe_ids <- joined_data %>%
  filter(duplicated(sample_id_key)) %>%
  pull(sample_id_key)

joined_data %>%
  filter(sample_id_key %in% dupe_ids ) %>%
  select(LUAD_soulette_barcode=barcode,
         gdc_sample_id = sample_id) %>%
 # distinct() %>%
  arrange(LUAD_soulette_barcode)
         
  
```

# format
colnames in manifest should be id      filename        md5     size    state

Find differences in gdc data between samples with the same barcode
```{r}


all_data_for_one_barcode <- joined_data %>%
  filter(barcode == "TCGA-44-2656-01A-02R-A278-07") %>%
  mutate(across(everything(), as.character)) %>%
  pivot_longer(-file_id) %>%
  pivot_wider (names_from = "file_id") 

all_data_for_one_barcode

all_data_for_one_barcode %>%
  filter(`b2cf0b26-1f46-478d-b2e1-73115384ed57` != `b5310007-7004-4306-8629-4c5a61219971`)
# 
# joined_data %>%
#   select(id, filename, md5, size, state)
# 
# joined_data %>%
#   filter(barcode == "TCGA-44-2656-01A-02R-A278-07") %>%
#   distinct()

```
there is no apparent difference besides file_id and file_name. the file sizes are identical

decision: pick at random



# pick at random
```{r}
  

set.seed(1)

data_for_new_manifest <- joined_data %>%
  group_by(barcode) %>%
  slice_sample(n=1)
  
nrow(data_for_new_manifest)

```

propogate to manifest

```{r}

soulette_equivalent_manifest <- gdc_manifest %>%
  filter(filename %in% data_for_new_manifest$file_name)

```

```{r}

write_tsv(soulette_equivalent_manifest, "soulette_equivalent_manifest.2026.05.28.tsv")

```

